---
title: "SummaryWidgets"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: yeti
---

```{r setup, include=FALSE}
library(flexdashboard)
library(SummaryWidgets)
library(crosstalk)
library(reactable)
library(bslib)
```


```{r}

shared_mtcars = SharedData$new(mtcars[,c(1,2,4,6,7,9)])

count_rows <- summaryTextSpan(
  data = shared_mtcars,
  statistic = "count")

average_mpg <- summaryTextSpan(
  data = shared_mtcars,
  statistic = "mean",
  column = "mpg",
  digits = 1,
  unit = "mile-per-gallon") #well formed unit https://tc39.es/ecma402/#sec-iswellformedunitidentifier

pct_automatic <- summaryTextSpan(
  data = shared_mtcars,
  statistic = "pct_total",
  selection = ~am==0,
  digits = 0, ## when digits == 0 0.1234 -> 12%
  number_format = "percent") 


pct_automatic_vb <- summaryValueBox(
  data = shared_mtcars,
  statistic = "pct_total",
  selection = ~am==0,
  digits = 0, ## when digits == 0 0.1234 -> 12%
  number_format = "percent",
  icon = "geo")


dynamic_caption <- bquote(
  sprintf("My caption is dynamic: %s", 
          summaryTextSpan(
            data = shared_mtcars,
            statistic = "count")))

count_rows_vb <- summaryValueBox(
  data = shared_mtcars,
  statistic = "count",
  caption = "let's test other things for now",
  color_thresholds = list(domain = c(0),
                          range = c("red","gold")))

average_mpg_vb <- summaryValueBox(
  data = shared_mtcars,
  statistic = "quantile",
  column = "mpg",
  digits = 1,
  unit = "mile-per-gallon",
  quantile = 0.5,
  caption = "Showing median value, can be adapted to other quantiles",
  icon = "heart",
  color_thresholds = list(domain = c(20,25),
                          range = c("#6a0032","#036064","#8b9538"))
  ) #well formed unit https://tc39.es/ecma402/#sec-iswellformedunitidentifier


pct_automatic2 <- summaryGauge(
  data = shared_mtcars,
  statistic = "pct_total",
  selection = ~am==0,
  digits = 0, ## when digits == 0 0.1234 -> 12%
  min = 0.33,
  max = 0.66,
  number_format = "percent")

pct_gears_automatic <- summaryGauge(
  data = shared_mtcars,
  statistic = "sum_pct_total",
  column = "hp",
  selection = ~am==0,
  digits = 0, ## when digits == 0 0.1234 -> 12%
  number_format = "percent")

maxQSec <- summaryGauge(
  data = shared_mtcars,
  statistic = "max",
  column = "qsec",
  min = 0,
  max = 50,
  digits = 1,
  unit = "second")
```

Column {data-width=250}
-----------------------------------------------------------------------

### Filter options
```{r example, echo=FALSE}
filter_checkbox("cyl", "Cylinders", shared_mtcars, ~cyl, inline = TRUE)
filter_slider("hp", "Horsepower", shared_mtcars, ~hp, width = "100%")
```


### Value Box 1: % automatic cars, icon, no caption

```{r}

pct_automatic_vb

```

### Value Box 2: Number of rows, no icon, caption

```{r}

count_rows_vb

```

### Value Box 3: Average MPG, icon, long caption

```{r}

average_mpg_vb

```

Column
-----------------------------------------------------------------------

### Summary widget in R markdown text

You have selected `r count_rows` cars with an average mpg of
`r average_mpg`. Of the selected cars, `r pct_automatic` are automatic.

### Summary widget in guages, wrapped in bscols (% automatic and max Qsec)

```{r}

bscols(pct_automatic2, maxQSec)

```

### Percentage of total horsepower belonging to automatic cars

```{r}

pct_gears_automatic

```

### Summary widget in guages check 2 (maxQSec, with dynamic min/max)

```{r}

maxQSec

```


Column {data-width=650}
-----------------------------------------------------------------------

### Summary widget as the footer in a reactable

```{r}

reactable(
  shared_mtcars,
  defaultColDef = colDef(
    align = "center",
    headerStyle = list(background="white"),
    footerStyle = list(fontWeight="bold")
  ),
  columns = list(
   mpg = colDef(
     name ="Miles per gallon",
     footer =  summaryTextSpan(
       data = shared_mtcars,
       statistic = "mean",
       column = "mpg",
       digits = 1,
       unit = "mile-per-gallon")
     ),
   wt = colDef(
     name = "Weight (x1000)",
     footer =  summaryTextSpan(
       data = shared_mtcars,
       statistic = "sum",
       column = "wt",
       digits = 0,
       unit = "pound") 
       ),
   am = colDef(
     name ="Automatic? (0 = Automatic)",
     footer = summaryTextSpan(
       data = shared_mtcars,
       statistic = "pct_total",
       selection = ~am==0,
       digits = 0,
       number_format = "percent") 
   )
  ),
  bordered = TRUE,
  striped = TRUE,
  highlight = TRUE)
```

